---
author: "Your Name"
title: "Computer Project 2: Probability"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Flipping Coins

For a fair coin, the chance of getting tails is 1/2.

## Counting the number of tails

Let's experiment with the total number of tails in $n$ trials.

###  Total number of tails in a small number of coin flips

Suppose we flip 10 coins (with 1 = Tails and 0 = Heads).
Here are six samples of size 10.

```{r}
n      = 10
trials = 6
for (i in 1:trials) {
   print(sample(c(0, 1), n, replace=TRUE))
}
```

Let's count the number of tails then subtract the expected number of tails.  The expected number
of tails is half the number of coin flips. 
```{r}
for (i in 1:trials) {
   print(sum(sample(c(0, 1), n, replace=TRUE))  -  n/2)
}
```

Based on the calculations, approximately what is the $\pm$ on the difference between
the total number of tails and  the expected total? (Remember:  95\% of the area under
a normal curve is $\pm 2$SD.)  In other words, approximately what is the SD of the list
of numbers shown above?

_Include your answer here._

\newpage

### Total number of tails in a large number of coin flips


Let's repeat the experiment with 10,000 coins.
```{r}
n = 10000
for (i in 1:trials) {
   print(sum(sample(c(0, 1), n, replace=TRUE))  -  n/2)
}
```

Based on the calculations, approximately what is the $\pm$ on the difference between
the total number of tails and  expected total?

_Include your answer here._


Does the $\pm$ on the number of tails increase or decrease with the number of trails?
Briefly justify your conclusion.

_Include your answer here._

\newpage

## Fraction of tails

Now let's look at the fraction of tails in $n$ trials.

###  Fraction of tails in a small number of coin flips

Suppose we flip 10 coins and compute the fraction of tails (with 1 = Tails and 0 = Heads).
```{r}
n      = 10
trials = 6
for (i in 1:trials) {
   print(sum(sample(c(0, 1), n, replace=TRUE)) / n)
}
```

The expected fraction of  tails is 0.5.
Let's compute the difference between the observed fraction and the expected fraction:
```{r}
n      = 10
trials = 6
for (i in 1:trials) {
   print(sum(sample(c(0, 1), n, replace=TRUE)) / n  - 0.5)
}
```


Based on the calculations, approximately what is the $\pm$ on the difference between
the observed fraction of tails and the expected fraction (0.5)?  In other words, approximately
what is the SD of the list of numbers shown above?

_Include your answer here._

\newpage

### Fraction of tails in a large number of coin flips


The expected fraction of tails is 0.5.  Let's repeat the experiment
with a larger number of coins.
```{r}
n = 10000
for (i in 1:trials) {
   print(sum(sample(c(0, 1), n, replace=TRUE)) / n  - 0.5)
}
```

Based on the calculations, approximately what is the $\pm$ on the difference between
the observed fraction of tails and 0.5

_Include your answer here._


Does the $\pm$ on the difference between the observed fraction of tails and 0.5
increase or decrease with the number of trails? Briefly justify your conclusion.

_Include your answer here._


\newpage

# Binomial Formula

If we roll a die, the chance of getting a 4 is $1/6$.  If we roll 60 dice
we expect to get about ten 4s. The number of ways of getting exactly ten 4s out
of 60 rolls is:

```{r}
n = 60
k = 10
choose(n, k)
```


The chance of getting exactly ten 4s is given by the binomial formula:

```{r}
p = 1/6
dbinom(k, size=n, prob=p)
```


The chance of getting ten or fewer 4s is:

```{r}
pbinom(k, size=n, prob=p)
```

Why do the  two probabilities differ so much?

_Include your answer here._

