---
author: "Your Name"
title: "Computer Project 1: Regression"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Predicting Pollution Levels from Daily High Temperature

The airquality data set has daily air quality measurements in New York, May to 
September 1973.  Here is a plot of ozone levels (in ppb) as a function of 
daily high temperature (degrees F).

```{r}
plot(airquality$Temp, airquality$Ozone, col='blue', xlab="Temperature", ylab="Pollution")
```

_Based on the graph, would you say the correlation is positive or negative? Is the magnitude
close to 1, 0 or something in between?_

\newpage

## Correlation

Here is the correlation.

```{r}
cor(airquality$Temp, airquality$Ozone, use='complete.obs')
```

_What does the correlation tell you about the association between temperature
and ozone?  Is it positive or negative?  Is it a strong or weak association?_

\newpage

## Regression

Here is a calculation of the regression line.
```{r}
model <- lm(airquality$Ozone ~ airquality$Temp)
summary(model)
```

Now plot the data with the regression line.

```{r}
plot(airquality$Temp, airquality$Ozone, col='blue', xlab="Temperature", ylab="Pollution")
abline(model, col='green')
```

_Use the output to write down the equation for the regression line.
Does the data seem to fit the line well or would it be better to use a quadratic or some other
curve?_

\newpage

# Predicting Pollution Levels from Wind Speed

Here is a plot of ozone levels (in ppb) as a function of 
wind speed (mph).

```{r}
plot(airquality$Wind, airquality$Ozone, col='blue', xlab="Wind Speed", ylab="Pollution")
```

_Based on the graph, would you say the correlation is positive or negative? Is the magnitude
close to 1, 0 or something in between?_

\newpage

## Correlation

Here is the correlation.

```{r}
cor(airquality$Wind, airquality$Ozone, use='complete.obs')
```

_What does the correlation tell you about the association between wind speed
and ozone?  Is it positive or negative?  Is it a strong or weak association?_

\newpage

## Regression

Here is a calculation of the regression line.
```{r}
model <- lm(airquality$Ozone ~ airquality$Wind)
summary(model)
```

Now plot the data with the regression line.

```{r}
plot(airquality$Wind, airquality$Ozone, col='blue', xlab="Wind Speed", ylab="Pollution")
abline(model, col='green')
```

_Use the output to write down the equation for the regression line.
Does the data seem to fit the line well or would it be better to use a quadratic or some other
curve?_

\newpage

# Predicting Pollution Levels from Solar Radiation

Here is a plot of ozone levels (in ppb) as a function of 
solar radiation level (in lang).

```{r}
plot(airquality$Solar.R, airquality$Ozone, col='blue', xlab="Solar Radiation", ylab="Pollution")
```

_Based on the graph, would you say the correlation is positive or negative? Is the magnitude
close to 1, 0 or something in between?_

\newpage

## Correlation

Here is the correlation.

```{r}
cor(airquality$Solar.R, airquality$Ozone, use='complete.obs')
```

_What does the correlation tell you about the association between solar radiation
and ozone?  Is it positive or negative?  Is it a strong or weak association?_

\newpage

## Regression

Here is a calculation of the regression line.
```{r}
model <- lm(airquality$Ozone ~ airquality$Solar.R)
summary(model)
```

Now plot the data with the regression line.

```{r}
plot(airquality$Solar.R, airquality$Ozone, col='blue', xlab="Solar Radiation", ylab="Pollution")
abline(model, col='green')
```

_Use the output to write down the equation for the regression line.
Does the data seem to fit the line well or would it be better to use a quadratic or some other
curve? Which of the three variables would be the most helpful in predicting ozone levels?  
Which the least?  Explain your conclusions._

