---
author: "Your Name"
title: "Computer Project 4: Hypothesis Testing"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(strtoi(substr(Sys.getenv("USER"), 1, 1), base=36))
secret.mean = 48.5
secret.sd   = 5
```

# $z$-Test Examples

We think the average of the contents of a box is 50.
Let's take some samples then test the hypothesis.

## Sample Size = 10

Here are the results of a sample of size 10.

```{r}
n = 10
x = rnorm(n, mean=secret.mean, sd=secret.sd)
x
```

Compute the mean of our sample:
```{r}
m = mean(x)
m
```

and the SD of the sample:

```{r}
s = sd(x) * sqrt((length(x)-1)/length(x))
s
```

Let's run a z-test to see if `r m` is close enough to 50.
First we need the SE for the average:

```{r}
se = s / sqrt(n)
se
```

Now compute the z-score:

```{r}
z = (m - 50)/se
z
```

The $p$-value is the chance of getting a z-score that is as extreme or more
extreme than the observed value.  _Use either pnorm(z) or 1 - pnorm(z) below
to compute the p-value:_

```{r}
# Use pnorm(z) or 1 - pnorm(z) here
```

The null hypothesis is that the difference between `r m` and 50 is due to chance.
_What is the alternative hypothesis?  Do the data support the null hypothesis or the alternative?_

\newpage


## Sample Size = 100

Here are the results of a sample of size 100

```{r}
n = 100
x = rnorm(n, mean=secret.mean, sd=secret.sd)
x
```

Compute the mean of our sample:
```{r}
m = mean(x)
m
```

and the SD of the sample:

```{r}
s = sd(x) * sqrt((length(x)-1)/length(x))
s
```

Let's run a z-test to see if `r m` is close enough to 50.
First we need the SE for the average:

```{r}
se = s / sqrt(n)
se
```

Now compute the z-score:

```{r}
z = (m - 50)/se
z
```

The $p$-value is the chance of getting a z-score that is as extreme or more
extreme than the observed value.  _Use either pnorm(z) or 1 - pnorm(z) below
to compute the p-value:_

```{r}
# Use pnorm(z) or 1 - pnorm(z) here
```

The null hypothesis is that the difference between `r m` and 50 is due to chance.
_What is the alternative hypothesis? Do the data support the null hypothesis or the alternative?_

\newpage

# t-Test Example

R has a built-in t-test command.  Let's use it to run the $n=100$ hypothesis test that
we did above.

```{r}
t.test(x, mu=50)
```

The output gives us the t-score, the degrees of freedom, the p-value and a 95%
confidence interval.

_Use the p-value to explain if the data supports the null or alternative hypothesis.
Is 50 included in the confidence interval?_













