---
title: "Report_4"
author: "Your Name"
date: "9 December 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Instructions:  

1. Update the read.csv command so that R knows where to find the data on your computer. 
2. Set the random number seed (see below) to a unique value.
3. Replace all the italic text with your own discussion (this is a focus of this report).
4. After you have completed the first confidence interval and z-test examples, repeat the analysis for the other cases described below.


## Introduction
The NHANES data set consists of four data files:  adult.csv, youth.csv, lab.csv and 
exam.csv.  The adult file contains information about subjects who were over 17 years
old.  The youth file contains information about younger subjects.  The lab and exam
files contain additional data about both adults and youth.


### Reading the data into R
For this report, we only need to read the exam.csv file into R.
```{r}
exam = read.csv("../nhanes/exam.csv", header=TRUE)
```

The size of this data set is
```{r}
dim(exam)
```

There are `r dim(exam)[1]` rows and `r dim(exam)[2]` columns in the exam file.

### Setting a Random Number Seed.

_Change the 10 in the following command to some other positive whole number.  Pick a number that is different from that of other students.  This value is used to randomly select a subject from the data set._
```{r}
set.seed(10)
```

\newpage
# Confidence Intervals

First extract the data that we will use for the confidence interval and z-test
examples:
```{r}
x1 <- data.frame(SEQN=exam$SEQN, HSSEX=exam$HSSEX, HSAGEIR=exam$HSAGEIR,  
                 PEP8=exam$PEP8, PEPMNK1R=exam$PEPMNK1R)
x2 <- na.omit(x1)
x3 <- x2[x2$PEP8 != 8,  ]
x4 <- x3[x3$PEPMNK1R != 888, ]
```

The size of this data set is
```{r}
dim(x4)
```

There are `r dim(x4)[1]` rows and `r dim(x4)[2]` columns in the exam file.

## Confidence Interval on Age (sample size = 25)

What is the average age of the subjects?  We can use a confidence interval 
to get an estimate from a small sample.

```{r}
n = 25
N = dim(x4)[1]
sample_rows = sample(1:N, n)
sample_data = x4[sample_rows, ]
```

Now compute the mean and sd of the ages in the sample:
```{r}
m = mean(sample_data$HSAGEIR)
m
s = sd(sample_data$HSAGEIR)
s
```

Compute the standard error (for the average):
```{r}
se = s / sqrt(n)
se
```

The 95% confidence interval on the true average age is 
from `r m - 2 * se` to `r m + 2 * se`.  The true average age of all subjects
is `r mean(x4$HSAGEIR)`.


\newpage

## Confidence Interval on Age (sample size = 100)

_Repeat the previous example using n = 100._

\newpage

# z-Tests

If the sample size is large, we use the z-test to test the null hypothesis
that the mean of a sample differs from a hypothesized mean simply due to 
chance.

## Blood Pressure of a Sample of Male Subjects

We can find the average systolic blood pressure, take a sample of the 
male subjects, then test the hypothesis that the average blood pressure
for the males is equal to the population average.


Overall, the average K1, systolic blood pressure of all subjects was:
```{r}
m = mean(x4$PEPMNK1R)
m
```

Extract the male subjects from the data set:
```{r}
men = x4[x4$HSSEX == 1, ]
```

Randomly select a sample of the male subjects.
```{r}
n = 100
N = dim(men)[1]
men_sample_rows = sample(1:N, n)
men_sample = men[men_sample_rows, ]
```

Compute the mean, standard deviation and standard deviation (for the average) 
of blood pressure measurements 
for this sample:
```{r}
x = mean(men_sample$PEPMNK1R)
s = sd(men_sample$PEPMNK1R)
se = s / sqrt(n)
```

Based on this sample, test the hypothesis that the blood pressure of 
the male subjects was equal to the blood pressure for all subjects.
The null hypothesis is that the difference between the overall average
and the sample average is due to chance.  The alternative hypothesis is 
that the difference is not due to chance.
```{r}
z = (x - m) / se
z
```

By the Central Limit Theorem, the sample mean is approximately a normal
distribution.  So we can compute the p-value as follows.
```{r}
1 - pnorm(z)
```
Since this p-value is small, it supports the alternative hypothesis.  The average male blood pressure seems to differ (be higher than) the overall average.

\newpage

## Blood Pressure of a Sample of Female Subjects

_Repeat the above example using just the female subjects.  Start with the line
women = x4[ ... ]_

\newpage

## Blood Pressure of a Sample of Subjects of a Specified Age

_Repeat the above example using just the subjects of a specified age.  Start with the line
subjects = x4[ ... ]_
