---
title: "Report_1"
author: "Me"
date: "23 September 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading the data into R
The NHANES consists of four data files:  adult.csv, youth.csv, lab.csv and 
exam.csv.  The adult file contains information about subjects who were over
17 years old.  The youth file contains information about other subjects.  The
lab and exam files contain additional data about both adults and youth.

### Reading the adult.csv file
The following command reads the adult.csv data file into R.
```{r}
adult = read.csv("../nhanes/adult.csv", header=TRUE)
```
The size of this data set is
```{r}
dim(adult)
```
This means that the file has 20,050 rows and 1238 columns.  So, there are
20,050 adults in the data set and 1238 pieces of information about each in 
this file.  The lab.csv and exam.csv include additional
information about these subjects.

### Reading the exam.csv file
The following command reads the exam.csv data file into R.
```{r}
exam = read.csv("../nhanes/exam.csv", header=TRUE)
```
The size of this data set is
```{r}
dim(exam)
```

## First Explorations
The first variable of interest is SEQN.  It stores the anonymous ID
numbers for the subjects.  Each subject has data stored in multiple files.
SEQN allows us to merge this disparate information for participants of
interest. Here are the first few ID numbers.
```{r}
head(adult$SEQN)
```
The following tells us that there is only one row in adult.csv for 
subject number 4.
```{r}
dim(adult[adult$SEQN == 4,])
```

### Variable DMARACER in the adult data set
The variable DMARACER gives information about the race of the subjects.
It is a qualitative, unordered variable but is recorded as a number:
1 = white, 2 = black, 3 = other and 8 = Mexican-American of unknown race.
We can confirm the distribution of DMARACER as follows:
```{r}
table(adult$DMARACER)
```

## Descriptive Statistics
There are two facets to statistics:  descriptive statistics and inferential
statistics.   Descriptive statistics involves organizing, summarizing and
visualizing data.  Inferential statistics involves making predictions.

### Variable BMPWT in the exam data file.
Subjects' body measurements were taken in a home exam.  Variable BMPWT
records subjects' weights in kg.  Some subjects did not have their weight
recorded, however.  The BMPWT variable is recorded as 888888 for these subjects.
So, if we compute the average weight without accounting for that fact:
```{r}
mean(exam$BMPWT)
```
we get 5135.477 kg.  That is a bit high for human patients.  We can extract
the true weights and compute some statistics as follows:
```{r}
e1 = exam[exam$BMPWT != 888888, ]
weights = e1$BMPWT
mean(weights)
sd(weights)
hist(weights, xlab="Weight (kg)", ylab="Percent per kg", probability=T)