---
title: "Central Limit Theorem"
author: 
date:
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Central Limit Theorem

The Central Limit Theorem says that:  When drawing with replacement from a
box, the probability histogram for the sum (and the average) will follow
the normal curve, even if the contents of the box do not.  The histogram
must be put into standard units, and the number of draws must be reasonably
large.

For a sum, the mean and standard deviation of this normal curve are:
\[\mbox{EV}_{\mbox{sum}} = n\cdot \mbox{AV}_{\mbox{box}}\]
and
\[\mbox{SE}_{\mbox{sum}} = \sqrt{n}\cdot \mbox{SD}_{\mbox{box}}\]
For an average, the mean and standard deviation of this normal curve are:
\[\mbox{EV}_{\mbox{av}} =\mbox{AV}_{\mbox{box}}\]
and
\[\mbox{SE}_{\mbox{av}} = \frac{\mbox{SD}_{\mbox{box}}}{\sqrt{n}}\]

The Central Limit Theorem is one reason that the normal curve plays such a central role in this course.

# Function Definitions
If you look in the Rmd file, you will see some functions defined in this
section.  They are used later in these notes to run simulations and make plots but are not printed here.
```{r echo=FALSE}
boxSimulation <- function(box, draws, repetitions) {
    y <- 1:repetitions;
    for (i in 1:repetitions)
      y[i] = sum(sample(box, draws, replace=TRUE))
    y;
  }
```

```{r echo=FALSE}
boxHistogram <- function(box, draws, repetitions, ymax=1.0, color="gray", breaks="Sturges", ylim=c(0, 1)) {
    y <- 1:repetitions;
        for (i in 1:repetitions)
          y[i] = sum(sample(box, draws, replace=TRUE))
        hist(y,
             prob=TRUE, # Generate a probability histogram 
             ylim=c(0, ymax), # Limits on the y axis
             ylab="",        # Label for the y axis
             main=paste(repetitions, "Repetitions"),  # Title for the graphic
             xlab=paste("Sum of ", draws, "draws from a box"), # x label
             col=color,      # Color for the bars
             breaks=breaks)  # Where to put breaks for the bars
}
```

```{r echo=FALSE}
diceSimulation <- function(repetitions, dice=2, diceSum=TRUE)
  {
    y <- 1:repetitions;
    if (diceSum == TRUE)
      for (i in 1:repetitions)
        y[i] = sum(sample(1:6, dice, replace=TRUE))
    else
      for (i in 1:repetitions)
        y[i] = prod(sample(1:6, dice, replace=TRUE));
    y;
}

diceHistogram <- function(repetitions, dice=2, ymax=0.3, diceSum=TRUE, color="gray")
  {
    y <- 1:repetitions;
    if (diceSum == TRUE)
      {
        for (i in 1:repetitions)
          y[i] = sum(sample(1:6, dice, replace=TRUE))
        hist(y, breaks=0.5+(1:(6*dice)), prob=TRUE,
             ylim=c(0,ymax), xlab=paste("Sum of Spots on", dice, "dice"),
             ylab="", col=color, main=paste(repetitions, "Repetitions"))
      }
    else
      {
        for (i in 1:repetitions)
          y[i] = prod(sample(1:6, dice, replace=TRUE));      
        hist(y, breaks=0.5+(0:(10+6^dice)), prob=TRUE,
             ylim=c(0,ymax), xlab=paste("Product of Spots on", dice, "dice"),
             ylab="", col=color, main=paste(repetitions, "Repetitions"))
      }
  }

```

Here is a function for computing the _population standard deviation_ of a box. It has $n$ in the denominator rather than $n-1$.  The _sample standard deviation_
uses $n-1$.
```{r}
SD <- function(box) {
  n = length(box)
  return (sd(box) * sqrt((n-1)/n))
}
```

\newpage

# Flipping Coins: and the Normal Curve
Flipping coins and counting the number of heads is like taking the sum
of a random sample (with replacement) from a 
box with two tickets 1 (for heads) and 0 (for tails).

## Five Coins:  Sum of the Number of Heads Follows a Normal Curve
The following simulates flipping five coins 100 times.
```{r}
data = boxSimulation(c(0, 1), 5, 100)
data
table(data)
```

Here are histograms of repetitions of this experiment:
```{r}
boxHistogram(c(0,1), 5, 100, breaks=-0.5+(0:6))
```
\newpage
```{r}
boxHistogram(c(0,1), 5, 100, breaks=-0.5+(0:6))
boxHistogram(c(0,1), 5, 100, breaks=-0.5+(0:6))
```

It looks a bit like a normal curve but changes from one sample to another.
Let's flip five coins 1000 times.
```{r}
boxHistogram(c(0,1), 5, 1000, breaks=-0.5+(0:6))
boxHistogram(c(0,1), 5, 1000, breaks=-0.5+(0:6))
boxHistogram(c(0,1), 5, 1000, breaks=-0.5+(0:6))
```

\newpage

## Ten Coins: Sum of the Number of Heads Follows a Normal Curve
The following simulates flipping five coins 100 times.
```{r}
data = boxSimulation(c(0, 1), 10, 100)
data
table(data)
```

Here are histograms of repetitions of this experiment:
```{r}
ymax=0.5
boxHistogram(c(0,1), 10, 100, breaks=-0.5+(0:11), ymax=ymax, col='blue')
```
\newpage
```{r}
boxHistogram(c(0,1), 10, 100, breaks=-0.5+(0:11), ymax=ymax, col='blue')
boxHistogram(c(0,1), 10, 100, breaks=-0.5+(0:11), ymax=ymax, col='blue')
```

It looks a bit like a normal curve but changes from one sample to another.
Let's flip 10 coins 10,000 times.
```{r}
boxHistogram(c(0,1), 10, 10000, breaks=-0.5+(0:11), ymax=ymax, col='blue')
boxHistogram(c(0,1), 10, 10000, breaks=-0.5+(0:11), ymax=ymax, col='blue')
boxHistogram(c(0,1), 10, 10000, breaks=-0.5+(0:11), ymax=ymax, col='blue')
```
\newpage

# Rolling Dice

We can simulate rolling dice and taking the sum )of the numbers shown.

## Two Dice: Sum of the Number of Dots Follows a Normal Curve
We could compute the frequencies using the 6 x 6 table of possible outcomes.
```{r}
data = diceSimulation(100)
data
table(data)
```

Let's look at histograms of 100 trials taking the sum of two dice.
```{r}
diceHistogram(100)
```
\newpage
```{r}
diceHistogram(100)
diceHistogram(100)
```

The results vary a bit each time we run 100 trials.  If we run 1000 trials, the 
results will be closer to the expected values.
```{r}
diceHistogram(1000)
diceHistogram(1000)
```

\newpage

## Three Dice:  Sum of the Number of Sots Follows a Normal Curve
Let's try the experiment with three dice.
```{r}
diceHistogram(10000, dice=3, color="green")
```
\newpage
```{r}
diceHistogram(10000, dice=3, color="green")
diceHistogram(10000, dice=3, color="green")
```

\newpage

# Sampling from a Box.

Some times it can take a very large sample size before the probability
histogram follows a normal curve very closely.

## Sample of Size 25 from {1, 2, 9} 
```{r}
boxHistogram(c(1,2,9), 25, 100, ymax=0.1, breaks=160, col='red')
```
\newpage
```{r}
boxHistogram(c(1,2,9), 25, 1000, ymax=0.1, breaks=160, col='red')
boxHistogram(c(1,2,9), 25, 10000, ymax=0.1, breaks=160, col='red')
boxHistogram(c(1,2,9), 25, 100000, ymax=0.1, breaks=160, col='red')
```

\newpage

## Sample of Size 50 from {1, 2, 9} 
```{r}
boxHistogram(c(1,2,9), 50, 1000, ymax=0.04, breaks=160, col='orange')
```
\newpage
```{r}
boxHistogram(c(1,2,9), 50, 10000, ymax=0.04, breaks=160, col='orange')
boxHistogram(c(1,2,9), 50, 100000, ymax=0.04, breaks=160, col='orange')
```

\newpage

# Product

The Central Limit Theorem says that when drawing with replacement from a
box, the probability histogram for the sum (and the average) will follow
the normal curve, even if the contents of the box do not.  It doesn't say
anything about the product or some other combination of the sample values.

## Product of Dice Rolls

If we roll two dice and take the product, we can get certain values
between $1=1\times 1$ and $36=6\times 6$.  Here is data generated
from 100 trials.
```{r}
diceSimulation(100, diceSum=FALSE)
```

Here are histograms from two such experiments:
```{r}
diceHistogram(1000, diceSum=FALSE, col='purple')
diceHistogram(10000, diceSum=FALSE, col='purple')
```

These don't look like normal curves at all.  Why not?

